Evolution of ORM Systems: Formal Foundations

### 2013-01-26 doc/description.tex
case study, updated insertValue

\paragraph{Insert Value} The function inserts a new pair or pairs into an existing row.
\begin{align*}
&	\mathbf{insertValue} : Row \times Pair* \times Database \rightarrow Daatabase \\
&	insertValue(r, p, d) = \begin{cases}
 		%\inference{|p| > 1}{\begin{gathered} 	d \bullet insertValue(r, head(p), d) \\ \bullet insertValue(r, tail(p), d)\end{gathered}}   \\ \\
 		\inference{|p| = 1 \wedge \forall e \in pairs(r) :\\  reference(e) \neq reference(p) \\ \wedge \exists t \in tables(d) : name(t) = refernces(r), \\ \wedge \exists c \in columns(t) : reference(p) = name(c)}{pairs(r) \cup p} 
 		%\\ \\ Inconsistent \; d
 		\end{cases} \\
& r = \{ x | x \in Row\}, p = \{ x | x \in Pair\},d \in Databases
\end{align*}

\paragraph{Add Value To Table} The function appends a pair to all rows of a table with given name.
\begin{align*}
&	\mathbf{addValueToTable} : Label \times Pair \times Database \rightarrow Database \\
&	addValueToTable(l, p, d) = \begin{cases}
 		\inference{\exists t \in tables(d) : name(t) = l}	{\begin{gathered}
		d \bullet insertValue(head(instances(l, d)), p, d) \\ \bullet addValueToTable(tail(instances(l, d)), p, d)
	\end{gathered}} 
 	\end{cases}\\ 
& 	l \in Label, p \in Pair,  d \in Database
\end{align*}

\paragraph{Copy Value} The function copy value from one column in a table to another column in the same or in another table - according to provided mapping. First argument of the transformation is the name of the source table, the second one is the target table, then there are names of source column and target column, finally the mapping between rows (stored instances) has to be provided.
\begin{align*}
&	\mathbf{copyValue} : Label \times Label \times Label \times Label \times (Row \times Row) \times Database \rightarrow Database \\
&	copyValue(l_s, l_t, l_{sc}, l_{tc} o_{l_s, l_t}, d) = \begin{cases}
		\inference{\exists t_s \in tables(d) : name(t_s) = l_s, \\ t_t \in tables(d) : name(t_t) = l_t, \\ c \in columns(t_s) : name(c) = l_{sc}, \\ e \in columns(t_t) : name(e) = l_{tc}}{\begin{gathered}
		\forall	p \in m_{l_s, l_t} :	d \bullet insertValue(second(p), \\ [l_{tc},value(findPair(l_{sc}, first(p)))])
		\end{gathered}}
 	\end{cases} \\
& 	l_s, l_t, l_{sc}, l_{tc} \in Label, d \in Database
\end{align*}

@@

\paragraph{Extract and Inline Class}
The extract and inline class are two opposite transformations. First of them extract a class from existing class, second of them merges two classes into one.
\begin{align*}
&	\mathbf{Extract} : Label \times Label \times Label \times (Row \times Row) \times State \rightarrow State \\
&	s \oplus Extract(l_s, l_n, l_p, m_{l_s, l_s}) = \begin{cases}
	\inference{|ps| = 1}{\begin{gathered}
		s \oplus addClass(cl(l_n, \varnothing, \varnothing)) \\ \oplus MoveProperty(l_s, l_n, ps, m_{l_s, l_s}) \end{gathered}}\\\\
	%\inference{|ps| > 1}{\begin{gathered} 	s \oplus Extract(l_s, head(ps), l_n) \\\oplus Extract(l_s, tail(ps), l_n) \end{gathered}}\\\\
	% \inference{|ps| = \varnothing}{state(s)}	 
 \end{cases}
	\\
&	l_s, l_n, l_p \in Label,  s \in State 
\end{align*}
The opposite transformation $Inline$ gets the name of a class to inline and the name of target class. The result of the transformation is a merge of information from two classes in the target class, the source class is removed. The important condition of feasibility is there is no name collision among properties in both classes.
\begin{align*}
&	\mathbf{Inline} : Label \times Label \times (Row \times Row) \times State \rightarrow State \\
&	s \oplus Inline(l_s, l_t, m_{l_s, l_t}) = \inference{\exists c_s \in classes(application(s)) : name(c_s) = l_s \\
\wedge \exists c_t \in classes(application(s)) : name(c_t) = l_t \\
\wedge \forall p_s \in properties(c_s), p_t \in properties(c_t) : name(ps) \neq name(p_t) \\
\wedge !\:associated(l_s, s)   }{\begin{gathered}
	s \oplus MoveProperty(l_s, l_t, ps, m_{l_s, l_t}) \\ \oplus RemoveClass(l_s) 
\end{gathered}
}\\
& l_s, l_t \in Label, m_{l_s, l_t} \in M, s \in State, \\ 
& ps = \{x | x \in \forall c \in classes(application(s)) : \forall p \in properries(c) : name(p))\}
\end{align*}

@@

ExtactClass = copyProperty("street"
\begin{align*}
a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
& pr(street, APPSTRING, \varnothing), \\
& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
& pr(street, APPSTRING, \varnothing), \\
& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
& cl("Address_tmp1", pr(street, APPSTRING, \varnothing) , \varnothing)
& )
\end{align*}
\begin{align*}
t &= db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
& tab(LegalPerson, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
& tab("Address_tmp1", pk("Id"), col(street, APPSTRING, \varnothing, \varnothing),\\
& \varnothing)) 
\end{align*}
\begin{align*}
r &= row("person", ["Id", 1], ["surname", "Griffin"], \\
& ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]), \\
& row("person", ["Id", 2], ["surname", "Doe"], \\
& ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]), \\
& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"]) \\
\end{align*}

>>>>

\paragraph{Insert Value} The function inserts a new pair or pairs into an existing row.
\begin{align*}
&	\mathbf{insertValue} : Label \times N \times Pair* \times Database \rightarrow Daatabase \\
&	insertValue(l, n, p, d) = \begin{cases}
 		%\inference{|p| > 1}{\begin{gathered} 	d \bullet insertValue(r, head(p), d) \\ \bullet insertValue(r, tail(p), d)\end{gathered}}   \\ \\
 		\inference{|p| = 1 \wedge 
 		\\ \wedge \exist r \in instances(l, d) : value(findPair("Id", r)) = n
 		\\ \wedge \exists t \in tables(d) : name(t) = l, 
 		\\ \wedge \exists c \in columns(t) : reference(p) = name(c) 
 		\\ \wedge \forall e \in pairs(r) :\\  reference(e) \neq reference(p) }{pairs(r) \cup p} 
 		%\\ \\ Inconsistent \; d
 		\end{cases} \\
& r = \{ x | x \in Row\}, p = \{ x | x \in Pair\},d \in Databases
\end{align*}

\paragraph{Add Value To Table} The function appends a pair to all rows of a table with given name.
\begin{align*}
&	\mathbf{addValueToTable} : Label \times Pair \times Database \rightarrow Database \\
&	addValueToTable(l, p, d) = \begin{cases}
 		\inference{\exists t \in tables(d) : name(t) = l}	{\begin{gathered}
		d \bullet (\forall r \in instances(l, d) \\ : insetValue(reference(r), keyValue(r), p) \\ \bullet addValueToTable(tail(instances(l, d)), p, d))
	\end{gathered}} 
 	\end{cases}\\ 
& 	l \in Label, p \in Pair,  d \in Database
\end{align*}

\paragraph{Copy Value} The function copy value from one column in a table to another column in the same or in another table - according to provided mapping. First argument of the transformation is the name of the source table, the second one is the target table, then there are names of source column and target column, finally the mapping between rows (stored instances) has to be provided.
\begin{align*}
&	\mathbf{copyValue} : Label \times Label \times Label \times Label \times (Row \times Row) \times Database \rightarrow Database \\
&	copyValue(l_s, l_t, l_{sc}, l_{tc} o_{l_s, l_t}, d) = \begin{cases}
		\inference{\exists t_s \in tables(d) : name(t_s) = l_s, \\ t_t \in tables(d) : name(t_t) = l_t, \\ c \in columns(t_s) : name(c) = l_{sc}, \\ e \in columns(t_t) : name(e) = l_{tc}}{\begin{gathered}
		\forall	p \in m_{l_s, l_t} :	d \bullet insertValue(name(second(p)), keyValue(second(p)), \\ [l_{tc},value(findPair(l_{sc}, first(p)))])
		\end{gathered}}
 	\end{cases} \\
& 	l_s, l_t, l_{sc}, l_{tc} \in Label, d \in Database
\end{align*}

@@

\paragraph{Extract and Inline Class}
The extract and inline class are two opposite transformations. First of them extract a class from existing class, second of them merges two classes into one.
\begin{align*}
&	\mathbf{Extract} : Label \times Label \times Label \times (Row \times Row) \times State \rightarrow State \\
&	s \oplus Extract(l_s, l_n, ps, m_{l_s, l_s}) = \begin{cases}
	\inference{|ps| = 1}{\begin{gathered}
		s \oplus addClass(cl(l_n, \varnothing, \varnothing)) \\ \oplus MoveProperty(l_s, l_n, ps, m_{l_s, l_s}) \end{gathered}}\\\\
	%\inference{|ps| > 1}{\begin{gathered} 	s \oplus Extract(l_s, head(ps), l_n) \\\oplus Extract(l_s, tail(ps), l_n) \end{gathered}}\\\\
	% \inference{|ps| = \varnothing}{state(s)}	 
 \end{cases}
	\\
&	l_s, l_n \in Label, ps ={ x | x  \in Label},  s \in State 
\end{align*}
The opposite transformation $Inline$ gets the name of a class to inline and the name of target class. The result of the transformation is a merge of information from two classes in the target class, the source class is removed. The important condition of feasibility is there is no name collision among properties in both classes.
\begin{align*}
&	\mathbf{Inline} : Label \times Label \times (Row \times Row) \times State \rightarrow State \\
&	s \oplus Inline(l_s, l_t, m_{l_s, l_t}) = \inference{\exists c_s \in classes(application(s)) : name(c_s) = l_s \\
\wedge \exists c_t \in classes(application(s)) : name(c_t) = l_t \\
\wedge \forall p_s \in properties(c_s), p_t \in properties(c_t) : name(ps) \neq name(p_t) \\
\wedge !\:associated(l_s, s)   }{\begin{gathered}
	s \oplus MoveProperty(l_s, l_t, ps, m_{l_s, l_t}) \\ \oplus RemoveClass(l_s) 
\end{gathered}
}\\
& l_s, l_t \in Label, m_{l_s, l_t} \in M, s \in State, \\ 
& ps = \{x | x \in \forall c \in classes(application(s)) : \forall p \in properries(c) : name(p))\}
\end{align*}

@@

\newpage
ExtactClass = copyProperty("street"
\begin{align*}
a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
& pr(street, APPSTRING, \varnothing), \\
& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
& pr(street, APPSTRING, \varnothing), \\
& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
& cl("Address_tmp1", pr(street, APPSTRING, \varnothing) , \varnothing)
& )
\end{align*}
\begin{align*}
t &= db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
& tab(LegalPerson, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
& tab("Address_tmp1", pk("Id"), col(street, APPSTRING, \varnothing, \varnothing),\\
& \varnothing)) 
\end{align*}
\begin{align*}
r &= \{ row("person", ["Id", 1], ["surname", "Griffin"], \\
& ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]), \\
& row("person", ["Id", 2], ["surname", "Doe"], \\
& ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]), \\
& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"]) \\
& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"] )\\
\}
\end{align*}

\newpage
ExtactClass = dropPropertyAndData("street"
\begin{align*}
a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
& pr(street, APPSTRING, \varnothing), \\
& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
& cl("Address_tmp1", pr(street, APPSTRING, \varnothing) , \varnothing)
& )
\end{align*}
\begin{align*}
t &= db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
& col(city, APPSTRING, \varnothing, \varnothing) \\
& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
& tab(LegalPerson, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
& tab("Address_tmp1", pk("Id"), col(street, APPSTRING, \varnothing, \varnothing),\\
& \varnothing)) 
\end{align*}
\begin{align*}
r &= \{ row("person", ["Id", 1], ["surname", "Griffin"], \\
& ["city", "NY"], ["zip", 10001]), \\
& row("person", ["Id", 2], ["surname", "Doe"], \\
& ["city", "LA"], ["zip", 90003]), \\
& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"]) \\
& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"] )\\
\}
\end{align*}

\newpage
ExtactClass = copyProperty("city"
\begin{align*}
a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
& pr(street, APPSTRING, \varnothing), \\
& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
& cl("Address_tmp1", pr(street, APPSTRING, \varnothing) pr(city, APPSTRING, \varnothing) , \varnothing)
& )
\end{align*}
\begin{align*}
t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \\
& col(city, DBSTRING, \varnothing, \varnothing) \\
& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
& tab("Address_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
& col(city, DBSTRING, \varnothing, \varnothing)
& \varnothing)) 
\end{align*}
\begin{align*}
r &= \{ row("person", ["Id", 1], ["surname", "Griffin"], \\
& ["city", "NY"], ["zip", 10001]), \\
& row("person", ["Id", 2], ["surname", "Doe"], \\
&  ["city", "LA"], ["zip", 90003]), \\
& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"]) \\
& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"], ["city", "LA"] )\\
\}
\end{align*}

\newpage
ExtactClass = dropPropertyAndData("city"
\begin{align*}
a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\
& pr(zip, APPSTRING, \varnothing)), \\
& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
& pr(street, APPSTRING, \varnothing), \\
& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
& cl("Address_tmp1", pr(street, APPSTRING, \varnothing) pr(city, APPSTRING, \varnothing) , \varnothing)
& )
\end{align*}
\begin{align*}
t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \\
& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
& tab("Address_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
& col(city, DBSTRING, \varnothing, \varnothing)
& \varnothing)) 
\end{align*}
\begin{align*}
r &= \{ row("person", ["Id", 1], ["surname", "Griffin"], \\
& ["zip", 10001]), \\
& row("person", ["Id", 2], ["surname", "Doe"], \\
& ["zip", 90003]), \\
& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"]) \\
& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"], ["city", "LA"] )\\
\}
\end{align*}

\newpage
ExtactClass = CopyProperty("zip"
\begin{align*}
a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\
& pr(zip, APPSTRING, \varnothing)), \\
& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
& pr(street, APPSTRING, \varnothing), \\
& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
& cl("Address_tmp1", pr(street, APPSTRING, \varnothing) pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing)
& )
\end{align*}
\begin{align*}
t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \\
& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
& tab("Address_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing)
& \varnothing)) 
\end{align*}
\begin{align*}
r &= \{ row("person", ["Id", 1], ["surname", "Griffin"], \\
& ["zip", 10001]), \\
& row("person", ["Id", 2], ["surname", "Doe"], \\
& ["zip", 90003]), \\
& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]) \\
& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]) )\\
\}
\end{align*}

\newpage
ExtactClass = DropPopertyAndData("zip"
\begin{align*}
a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing))), \\
& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
& pr(street, APPSTRING, \varnothing), \\
& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
& cl("Address_tmp1", pr(street, APPSTRING, \varnothing) pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing)
& )
\end{align*}
\begin{align*}
t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
& tab("Address_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing)
& \varnothing)) 
\end{align*}
\begin{align*}
r &= \{ row("person", ["Id", 1], ["surname", "Griffin"]), \\
& row("person", ["Id", 2], ["surname", "Doe"]), \\
& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]) \\
& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]) )\\
\}
\end{align*}

\newpage
$AddAssociation("Person", assoc("address", "Address_tmp1", 1, 1), 
 \{(x,y)| x \in instances("Person", database(S)), y \in instances("Address_tmp1", database(S)) : pkVal(x) = pkVal(y) \})$
\begin{align*}
a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing)), \\ 
& assoc("address", "Address_tmp1", 1, 1)), \\ 
& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
& pr(street, APPSTRING, \varnothing), \\
& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
& cl("Address_tmp1", pr(street, APPSTRING, \varnothing), \\
& pr(city, APPSTRING, \varnothing), pr(zip, APPSTRING, \varnothing), \varnothing)
& )
\end{align*}
\begin{align*}
t &= db(tab(Person, pk("Id"), col(surname, DBSTRING, \varnothing, \varnothing), \varnothing), \\
& tab(LegalPerson, pk("Id"),  col(businessName, DBSTRING, \varnothing, \varnothing), \\
& col(street, DBSTRING, \varnothing, \varnothing), col(city, DBSTRING, \varnothing, \varnothing) \\
& col(zip, DBSTRING, \varnothing, \varnothing), \varnothing), \\
& tab("Address_tmp1", pk("Id"), col(street, DBSTRING, \varnothing, \varnothing),\\
& col(city, DBSTRING, \varnothing, \varnothing), col(zip, DBSTRING, \varnothing, \varnothing)
& fk("address", "Person", \varnothing))\\
) 
\end{align*}
\begin{align*}
r &= \{ row("person", ["Id", 1], ["surname", "Griffin"]), \\
& row("person", ["Id", 2], ["surname", "Doe"]), \\
& row("legalPerson", ["Id", 1], ["businessName", "Tools and Machines ltd."], \\
& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
& row("legalPerson", ["Id", 2], ["businessName", "Doe"], \\
& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003])\\
& row("address\_tmp1", ["Id", 1], ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]) \\
& row("address\_tmp1", ["Id", 1], ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]) )\\
\}
\end{align*}
